---
title: "Code Bank: 2_basicApp_widgets"
author: "Emma Jones"
date: "4/30/2019"
output: html_document
---


## Background 

This script is meant to assist you both in and out of the workshop. Feel free to copy/paste code from this script into the appropriate files to run your app in case you miss a step as  we are walking through it together. You can also use this script as you build your own apps because it is organized in paired ui/server blocks of code that demonstrate complete reactive elements. 
 
We will apply the following steps in order to transform our ui/server files (ui_start.R and server_start.R, respectively) into the final app (final_app_scripts/ui_final.R and final_app_scripts/ui_final.R).


## Code

### Step 1: Fix the radiobutton server output such that it actually subsets data

#### SERVER
Put anywhere but preferably after the first two output$ functions for flow/understanding

Switch is critical for radio buttons and other user input widgets. It transforms the user input into a format you can use for your own data manipulation or server calls

```{r step1_server}

indicator <- reactive({
  switch(input$radio,"1"="VSCIAll","2"="TotalHabitat")})

cdfsubset <- reactive({
  cdf2 <- filter(cdfdata, Subpopulation == input$dropdown, Indicator == indicator() )
  return(cdf2)
})

```


### Step 2: Add a table to user interface to preview the data

#### UI
Put in mainPanel() after radioOutput call; DON'T FORGET COMMA!

```{r step2_ui}

tableOutput("tablePreview")

```

#### SERVER

Now display subsetted data to the user

```{r step2_server}

output$tablePreview <- renderTable({
  head(cdfsubset(),15)
})

```



### Step 3: Plot the cdfdata subset, add an option for user to toggle the confidence intervals on and off

#### UI
Put in the sidebarPanel after the radiobutton, DON'T FORGET COMMA!

```{r step3_ui.1}

hr(),
checkboxInput('conf',"Plot Confidence Intervals")

```


Put in the mainPanel after the tableOutput, DON'T FORGET COMMA!

```{r step3_ui.2}

br(),
br(),
hr(),
plotOutput('cdfplot')

```


#### SERVER

```{r step3_server}

output$cdfplot <- renderPlot({
  cdfplotFunction(cdfsubset(),input$dropdown,indicator(),input$conf)
})

```



### Step 4: Add a downloadButton to let users download a flat file of their data subset

#### UI

Put in sidebarPanel after checkboxInput, DON'T FORGET COMMA!

```{r step4_ui}

hr(),
br(),
downloadButton('downloadCDFsubset',"Download CDF data")

```



#### SERVER

Probably put this after all previous bits.

```{r step4_server}

 output$downloadCDFsubset <- downloadHandler(
    filename=function(){
      # filename, note we customized it based on the user subset data
      paste('cdfsubset',input$dropdown, indicator(), Sys.Date(),'.csv', sep='_')}, 
    
    # what would happen if you had just put in input$radio instead of indicator() ?
    
    content=function(file){
      write.csv(cdfsubset(),file)})

```



### Step 5: Add all Subpopulation options to selectInput()

#### UI

Delete selectInput() and put this instead (in sidebarPanel)

```{r step5_ui}

uiOutput('dropdownwithallcsvoptions'),

```



#### SERVER

Put anywhere but for flow make this the first function

Delete output$dropdownOutput, replace with this:

```{r step5_server.1}

output$dropdownwithallcsvoptions <- renderUI({
  selectInput("dropdownwithallcsvoptions","Select Population to Plot",
              unique(cdfdata$Subpopulation))})

```


Update the cdfsubset() accordingly

Because we added a reactive element to the server side, we need to give the server a break and let it process. Basically, insert the req() statement of the next chunk into cdfsubset() after the reactive({ call to prevent an error message from initially popping up when the app is first loaded. 

req() is a great function that tells the server to not do anything until whatever is inside the function either exists or if it changes. 


```{r step5_server.2}

# cdfsubset <- reactive({
  req(input$dropdownwithallcsvoptions)
# then the rest of that function

```


Because we gave cdfsubset() a second to think, we need to do the same for all of the subsequent functions that use cdfsubset()

```{r step5_server.3}

# output$tablePreview <- renderTable({
  req(cdfsubset())
# then the rest of that function

```


```{r step5_server.4}

#output$cdfplot <- renderPlot({
  req(cdfsubset())
# then the rest of that function

```

This will stop the server from initally freaking out that it doesn't have something to display/plot right when the app loads (because it is still processing the renderUI). This is a great trick for most reactive elements that give errors initially as the app is loading or if you don't want to display a table/plot/result until you first get something from a user. 

